\mytitle

### 2020-05-28 paper.tex
Eldan's comments.

Our experiments are focused on the performance of the two approaches
compared to 
existing optimization methods. To this end,
we experiment with 12 standard WFLO benchmark problems (c.f.~\cite{turner2014new})
and show that solving quadratic WFLO models
using Gurobi and the DA
achieves state-of-the-art performance.
%, often outperforming existing solutions when provided sufficient run-time.
Furthermore, we observe that the DA often provides solutions of
similar quality to software-based approaches but with two orders of
magnitude lower run-time.
%The ability to quickly obtain competitive
%solutions on large WFLO instances can be employed to solve multiple
%wind farm design problems.

@@

%\paragraph{Wind Farm as a Grid}
We consider a square wind farm that we model as an $n\times n$
grid with $n$ being the number of cells on each axis.  The cells have
an area of $c^2$ (in square meters), i.e., if $c=100$ meters and
$n=20$, the farm area will be $2km\times 2km$.  We
assume that the terrain is flat and that we have no constraints
related to the terrain.  Furthermore, we do not account for the
influence of noise on the wind farm's surroundings. The wind farm
literature treats the more complex farm models that capture non-flat
terrain~\cite{song2015lazy,kuo2016wind}, multi-typed
turbines~\cite{feng2017design}, noise
effects~\cite{Zhang2014,sorkhabi2016impact,yin2014multi}, and complex
objectives (including turbine installation and maintenance
costs)~\cite{lackner2007analytical,sorkhabi2018constrained}.  In this work, we focus on the
novel quadratic approach for WFLO, and hence consider the most
abstract wind farm model. Our model can be extended, in the
future, to include additional problem characteristics.

@@

\paragraph{Wakes}
Turbines influence each other through \emph{wake} interference effects~\cite{jensen1983note,shakoor2016wake}.
%that are
%referred to as \emph{wakes}~\cite{jensen1983note,shakoor2016wake}.
Wakes cause reduction of the effective power produced by a turbine due
to upstream turbines that change the airflow dynamics. In
Figure~\ref{fig:field_model}, the wind is assumed to be
blowing from left to right. The wake effect due to placing the turbine
expands as shown in the diagram.
%: it will influence only turbines that
%are placed in the orange cells.  In other words,
If a turbine is placed in the orange area, its effective power is
reduced due to the upstream turbine.
% that we placed in the $8\times8$ farm.
Moreover, a given turbine may be placed downstream from multiple turbines and hence
multiple wakes must be taken into account for each location.  The
reduction of power due to the wake effect is determined by three
parameters: (1) the distance between the upstream turbine and the cell
for which we want to compute the effect (denoted by $x$ in
Figure~\ref{fig:field_model}), (2) the radius of the cone's opening at distance $x$ 
(denoted by $r$ in Figure~\ref{fig:field_model}), and (3) the
combination of several wind regimes where a wind regime consists of 
the wind's free speed and direction.  The wake effect can be pre-computed
for each pair of cells since we know: the distance ($x$), the turbine
specifications and terrain conditions (that jointly dictate $r$), and
probabilistic behavior of wind regimes~\cite{Zhang2014}.

@@

Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$.\footnote{Turbines can adjust their orientation according to the current wind 
regime.}  Given the interference of the wakes, the velocity of turbine $i$ is not equivalent to the free wind speed of $d$.
Rather, $u_{id}$, can be computed using the following equation~\cite{Zhang2014}:
\begin{equation}
u_{id} = u_{id\infty} \Bigg[1 - \sqrt{\sum_{j\in\mathcal{U}_{id}} \bigg( 1-\frac{u_{ijd}}{u_{id\infty}} \bigg)^2}  \Bigg], \label{eq:ss}
\end{equation} with $u_{id\infty}$ being the free wind speed (the second component of regime $d$), $\mathcal{U}_{id}$ corresponding to the set of turbines  that are upstream to $i$ 
under wind regime $d$, and $u_{ijd}$ being the wake-reduced speed at $i$ due to upstream turbine $j$. 
The reduced speed, $u_{ijd}$, is computed using the distance $x$ between the two turbines ($i,j$)  and turbine and terrain specifications that yield the cone opening radius, $r$. The set of upstream turbines, $\mathcal{U}_{id}$ is computed  
by rotating the grid wrt to the current wind regime $d \in \mathcal{D}$ and identifying the turbines, $j$, located upstream of turbine $i$ given $d$. The expression in Eq. (\ref{eq:ss}) is referred to as the \emph{sum-of-squares} (SS) model for 
the total expected power in presence of wakes~\cite{Zhang2014}.  We can now write 
the sum-of-squares (SS) expected power of a wind farm with $m$ turbines:
%~(\cite{Zhang2014}):
\begin{equation}
  E_{SS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \frac{1}{3} u_{id}^3p_d.\label{eq:ess}
\end{equation}

@@

\paragraph{Exact Optimization Methods} 
The WFLO problem can be solved  
to optimality, using
techniques from operations research (e.g., mixed-integer linear programming (MILP), quadratic programming (QP), and constraint programming (CP))~\cite{Zhang2014,turner2014new,donovan2005wind}.
These methods are \emph{exact}, in the sense that they guarantee that the returned solution is indeed
the best solution one can obtain for the given problem. Donovan was the first to solve WFLO using a
MILP approach based on the LS energy expression~\cite{donovan2005wind}. Turner et al.\ \cite{turner2014new} approximated the SS energy expression using a quadratic and a linear approximation. The two resulting methods performed well when
compared to existing approximate and exact techniques.
Another non-linear optimization approach \cite{ulku2019new}, relaxed the binary placement variables to continuous values in $[0,1]$, 
and approximated the constraints such that the resulting variables yield an integer placement.
Their approach provides another approximation to the WFLO problem considered by Turner et al.\ \cite{turner2014new}.

@@

A benchmark study comparing state-of-the-art exact approaches \cite{Zhang2014} to 
approximate methods and found that the two perform in a comparable fashion, without large differences between
the attained energy~\cite{yang2019simulated}. Exact methods tend to perform slightly better
in smaller WFLO benchmarks, while simulated annealing methods dominated for larger standard WFLO benchmarks. 
These results lead to the conclusion that both types of methods (exact and approximate) should be considered. Therefore, we introduce a unified paradigm, namely quadratic programming, that enables
both exact and approximate solutions to WFLO. Further, formulating WFLO using quadratic programming enables us to employ 
recent advanced exact software solutions (e.g., Gurobi optimizer) and approximate hardware developments 
(e.g., Fujitsu's Digital Annealer).
%Thus, 
%our approach `enjoys' the best of both worlds (the exact and the approximate).

@@

%% In this section,
%% we formulate the WFLO problem
%% using quadratic programming. Specifically,
We consider two
quadratic optimization problems that represent WFLO: a
quadratic constrained optimization problem (QCOP) and 
a quadratic unconstrained binary optimization (QUBO) problem. 
The former represents WFLO characteristics (proximity and number of turbines)
as hard constraints, while the latter considers these characteristics as soft constraints that are penalized as part of the objective function.

@@

To 
maximize $f(x)$ and satisfy the total number of turbines and proximity 
constraints we write
the following quadratic optimization problem (QCOP):
\begin{eqnarray} \label{QCOP}
&\max_{x_i}^{}& \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)x_i x_j)   \\
&s.t.:& \nonumber\\
&\mbox{       }& \sum_{i \in \mathcal{N}}^{} x_i = m,\label{Cardinality}\\
&\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid}\\
&\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}. \label{Vars}
\end{eqnarray} Constraints~\ref{Cardinality}~and~\ref{Grid} ensure that exactly $m$ turbines are placed on the grid
and enforce proximity constraints between the turbines, respectively. We shall refer to the QCOP
formulation of WFLO as \qcls{}.

\qcls{} can be solved using an exact optimization solver, e.g., Gurobi~\cite{gurobi}. 
Similar, linear formulations have been previously proposed and solved using exact solvers~\cite{Zhang2014,donovan2005wind}.
%\todo{JCB: We need to formally present the ILP model in the experiment section. We can then discuss the fact that the model is identical - because the linearization gives and exact representation. Should put a forward reference to that model here.}
%\todo{Are these linear formulations ``equivalent'' as you claimed or approximations? If they are mathematically equivalent, why do we think a quadratic model will be any better computationally? Arik: The two formulations are equivalent to WFL-QCOP. However, we compare against the ILP model experimentally and it does worse than WFL-QCOP. Hopefully, Jason's part in Section 3.5.1. based on Wen-Yang's thesis will explain why it does better. Originally, I used the quadratic formulation because it seemed more natural. Semantically, the linear model and the quadratic model are the equivalent imho.}

@@

Let $\lambda = (\lambda_1,\lambda_2)$ 
be a vector of constraint parameters with $\lambda_i >0$. The equivalent QUBO formulation of the QCOP in Eq. (\ref{QCOP}) is given by,
\begin{equation}\max_{x}^{} f(x) - \lambda_1 (\sum_{i \in \mathcal{N}}^{} x_i -m) ^2 - \lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j , \label{QUBO}\end{equation}
with the term $ \lambda_1 (\sum_{i \in \mathcal{N}}^{} x_i -m) ^2$ 
penalizing solutions that violate the exactly $m$ turbines constraint and the term $\lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j$ penalizing
solutions that place turbines too close together. The two penalty terms $\lambda_1$ and $\lambda_2$ 
must be large enough to ensure that the solutions are indeed feasible wrt the problem constraints. We shall refer to the QUBO representation of WFLO as \quls.
%These types of constraints
%are referred to as \emph{soft constraints} as they do not guarantee a feasible solution 
%for an arbitrary selection of $\lambda$.

@@

\begin{mythm}
	The computational complexity of \qcls{} defined in \eqref{QCOP}-\eqref{Vars} is $\mathcal{NP}$-hard. 
\end{mythm}
\begin{myproof}
We prove by showing that a special case of the \qcls{} \eqref{QCOP}-\eqref{Vars}  is
$\mathcal{NP}$-hard. Suppose that $\mathcal{E} = \emptyset$, i.e., the grid is coarse grained enough to ignore proximity constraints. Then, the constraints in Eq.~(\ref{Grid})
are dropped and we get the heaviest k-subgraph problem, which was proven to be $\mathcal{NP}$-hard~\cite{billionnet2005different}.
Since a special case of \qcls{} is  $\mathcal{NP}$-hard it implies that \qcls{} is at least $\mathcal{NP}$-hard. On the other hand,
\qcls{} is formulated using integer programming (Problem \ref{QCOP}), which means that it is at most $\mathcal{NP}$-hard. Therefore, we get that \qcls{} is $\mathcal{NP}$-hard.
\end{myproof}

@@

In this section, we 
present an empirical evaluation of the two quadratic models based on 
12 standard WFLO instances.
Our main results are: \begin{itemize}
	\item \quls{} when solved using the Digital Annealer (DA) achieves a solution quality as good or better than the state-of-the-art exact methods on 10 of 12 instances.
		\item To achieve such a solution quality, the runtime for solving \quls{} on the DA is at least two orders of magnitude lower than \qcls{} for hard WFLO instances.  
	\item Given sufficient runtime, solving the \qcls{} model yields strong results on most of the instances. Compared to previous exact models (such as \qcss), \qcls{} is able to find equal or better solutions on 8 of 12 benchmark problems and for 5 of the 6 hard scenarios. %  \todo{Arik: Is this correct? Table 2 shows DA@10seconds with the best solution found for 7/12 instances and \qcls{}grb@3600 for 6/12. And 4/6 vs. 3/6 for the WR36). Are you comparing \qcls{}grb with \qcssgrb and ILP-LS? We seem to be underplaying DA performance in this comment. Chris: I changed to competitive results. You're right, I'd say that the DA is not `losing' to \qcls{}grb.}  
\end{itemize}

@@

Recall that $\mathcal{W}$ is the set of our 12 standard WFLO instances. 
Let $b_w$ be the best energy solution (in terms of SS) obtained for an instance $w \in \mathcal{W}$ 
across the different methods.
%For example, the best solution (within 3600 seconds) for the second instance $w_2$ that aims at placing $30$ turbines on a $10\times10$ grid with a single wind regime is attained by \qulsgrb and the SS energy expression is $15742$ (kW); thus, we set $b_w = 15742$.
Then, for every point in time $t \in \{0,\ldots, 3600\}$ (measured in seconds), we can compute the relative error of a given approach per instance $w$. Let $\mathcal{A}$ be the set of solution approaches, $$\mathcal{A}= \mbox{
	\{\emph{\qulsgrb}}, \mbox{\emph{\qclsgrb}}, \mbox{\emph{\qulsda}}, \mbox{\emph{\ilplsgrb}}, \mbox{\emph{\qcssgrb}}\}.$$ Further, denote by $b_{w,t,a}$ the best solution attained before time $t$ of approach $a$ for instance $w$. Then, the relative error at time $t$ for approach $a$ on instance $w$ is given by
\begin{equation}
  RE(w,t,a) = \frac{b_w - b_{w,t,a}}{b_w}.
\end{equation} Note that the expression is always positive, since we are solving a maximization problem. We can now average over the different instances at time $t$. The average performance of approach $a$ at time $t$, $MRE(t,a)$ can be computed as, \begin{equation} MRE(t,a) = \frac{1}{|\mathcal{W}|}\sum_{ w\in\mathcal{W}} RE(w,t,a).\end{equation}
We present $MRE(t,a)$ for each of the approaches to demonstrate the speed of convergence of the proposed LS-based methods.

@@

We observe that the DA-based solution provides
comparable results to state-of-the-art exact methods. For the harder instances (of $36$ wind directions),
the three quadratic LS-based approaches (\qulsda{}, \qulsgrb{}, and \qclsgrb{}) dominate the others by finding the best (or equal) solutions in 11 out of 12 instances. For easier instances, 
the linear approach (\ilplsgrb) finds better solutions.
%\todo{@Chris, Jason: do we have a good explanation for this?}. NO
Table~\ref{tab:results2} shows the same type of results as in Table~\ref{tab:results1}, but now after $3600$ seconds. For the exact technique, we can see that the improvement between 100 and 3600 seconds of run-time is marginal. We also observe that the DA at 10 seconds is, therefore, still competitive with the other approaches,
while \qclsgrb{} finds the best solutions in 10 out of the 12 scenarios. 
Our quadratic approaches dominate all six hard scenarios (12 wind directions). %\todo{@Chris: I am not sure that the first table contributes anything beyond the second table. JCB: ``We include the second table to demonstrate that only minor improvements are achieved by the exact techniques between 100 seconds of run-time.''}

We now turn to present the mean relative error over time per approach. Note that lower values indicate better performance.  Figure~\ref{fig:results1} 
presents the MRE after $50$ seconds, while Figure~\ref{fig:results2} corresponds
to the full run (after $3600$ seconds).
%The latter provides insights into best performance in terms of energy. The 5 different lines (and their colors) 
%correspond to the 5 WFLO solutions. The horizontal axis represents time in seconds.

Our proposed quadratic models, \qcls{} and \quls{},
outperform the rest of the techniques both in terms of runtime (best results in short time), and in terms of overall performance.
Specifically, for most instances, \qclsgrb{} gives the best solution at 3600 seconds, while the DA-based approach
yields competitive solutions after less than 10 seconds. 
Furthermore, 
\qclsgrb{} converges to the best solution faster than all other exact methods. In fact, it is inferior only to the approximate \qulsda{} method.

@@

\begin{sloppypar}To demonstrate the gain in runtime from using the DA, we compare the average speed-up factor between
 \qclsgrb{} and a \qulsda{} solutions after $1$ and $10$ seconds.
For each $w \in \mathcal{W}$ we computed the mean $S(\mbox{\emph{\qclsgrb}}, \mbox{\emph{\qulsda}}, t, w)$ for $t \in \{1,10\}$. 
%and $S(\qclsgrb, \qulsda, 10sec, w)$
%for every $w$ 
%and averaged the speed-up factors across the instances. 
The average speed-up factors were found to be $666$ and $100$ for $1$ and $10$ seconds, respectively. 
Per instance, the minimum speedup was $1$ for $1$ second and $0.1$ for $10$ seconds: for some instances \qclsgrb{} achieved the same quality as \qulsda{} in less time. The maximum speedup was (at least) $3600$ for $1$ second and (at least) $360$ for $10$ seconds: for those instances
\qclsgrb{} was not able to find a comparable solution to \qulsda{} within the $3600$ time limit. As a consequence, 
the \qulsda{} speed-up is actually even higher than computed above. \end{sloppypar}
%\footnote{The speed-up factors  were rounded to have 0 values after the floating point.}

%These results show that the DA achieves competitive solutions after running for $1$ second.
%Many of the instances running \qclsgrb for 3600 seconds did not result in as good a solution as \qulsda at time $t$ (1 or 10 seconds).
%we consider the speed-up factor to be a lower bound, meaning
When compared
to the other LS-based methods, the DA speed-up is even larger than when compared to \qclsgrb{}.  At $1$ ($10$) seconds, the speedup factor compared to \ilplsgrb{} 
is $1385$ ($180$) and for the \qulsgrb{} it is $900$ ($152$). We compare against the LS based methods as they directly optimize on the same objective as the DA.  
For full speedup results see Tables~\ref{tab:results3} and~\ref{tab:results4} in the Section~\ref{app1}.
%speed-up factor, which we did not compute due to the time limit of $3600$. 
%This is an encouraging result, especially due to the pessimistic computational complexity of WFLO.

>>>>

Our experiments are focused on the performance of the two approaches
compared to 
existing optimization methods. To this end,
we experiment with 12 standard WFLO benchmark problems (c.f.~\cite{turner2014new})
and show that solving quadratic WFLO models
using Gurobi and the DA
achieves state-of-the-art performance.
%, often outperforming existing solutions when provided sufficient run-time.
Furthermore, we observe that the DA often provides solutions of
similar quality to software-based approaches but with two orders of
magnitude lower \todo{Eldan: shorter instead of lower?} run-time.
%The ability to quickly obtain competitive
%solutions on large WFLO instances can be employed to solve multiple
%wind farm design problems.

@@

%\paragraph{Wind Farm as a Grid}
We consider a square wind farm that we model as an $n\times n$
grid with $n$ being the number of cells on each axis.  The cells have
an area of $c^2$ (in square meters), i.e., if $c=100$ meters and
$n=20$, the farm area will be $2km\times 2km$.  We
assume that the terrain is flat and that we have no constraints
related to the terrain.  Furthermore, we do not account for the
influence of noise on the wind farm's surroundings. The wind farm
literature treats the more complex farm models that capture non-flat
terrain~\cite{song2015lazy,kuo2016wind}, multi-typed
turbines~\cite{feng2017design}, noise
effects~\cite{Zhang2014,sorkhabi2016impact,yin2014multi}, and complex
objectives (including turbine installation and maintenance
costs)~\cite{lackner2007analytical,sorkhabi2018constrained}.  In this work, we 
choose to focus on a
novel approach for solving WFLO, and hence consider the most
abstract wind farm model. Our model can be extended, in the
future, to include additional problem characteristics as mentioned above.

@@

\paragraph{Wakes}
Turbines influence each other through \emph{wake} interference effects~\cite{jensen1983note,shakoor2016wake}.
%that are
%referred to as \emph{wakes}~\cite{jensen1983note,shakoor2016wake}.
Wakes cause reduction of the effective power produced by a turbine due
to upstream turbines that change the airflow dynamics. In
Figure~\ref{fig:field_model}, the wind is assumed to be
blowing from left to right. The wake effect due to placing the turbine
expands as shown in the diagram.
%: it will influence only turbines that
%are placed in the orange cells.  In other words,
If a turbine is placed in the orange area, its effective power is
reduced due to the upstream turbine.
% that we placed in the $8\times8$ farm.
Moreover, a given turbine may be placed downstream from multiple turbines and hence
multiple wakes must be taken into account for each location.  The
reduction of power due to the wake effect is determined by three
parameters: (1) the distance between the upstream turbine and the cell
for which we want to compute the effect (denoted by $x$ in
Figure~\ref{fig:field_model}), (2) the radius of the cone's opening at distance $x$ 
(denoted by $r$ in Figure~\ref{fig:field_model}), and (3) the
combination of several wind regimes where a wind regime consists of 
the wind's free speed and direction.  The wake effect can be pre-computed
for each pair of cells since we know the distance ($x$), the turbine
specifications and terrain conditions (that jointly dictate $r$), and the
probabilistic behavior of wind regimes~\cite{Zhang2014}.

@@

Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$.\footnote{Turbines can adjust their orientation according to the current wind 
regime.}  Given the interference of the wakes, the velocity of turbine $i$ is not equivalent to the free wind speed of $d$.
Rather, $u_{id}$ can be computed using the following equation~\cite{Zhang2014}:
\begin{equation}
u_{id} = u_{id\infty} \Bigg[1 - \sqrt{\sum_{j\in\mathcal{U}_{id}} \bigg( 1-\frac{u_{ijd}}{u_{id\infty}} \bigg)^2}  \Bigg], \label{eq:ss}
\end{equation} with $u_{id\infty}$ being the free wind speed (the second component of regime $d$), $\mathcal{U}_{id}$ corresponding to the set of turbines  that are upstream to $i$ 
under wind regime $d$, and $u_{ijd}$ being the wake-reduced speed at $i$ due to upstream turbine $j$. 
The reduced speed, $u_{ijd}$, is computed using the distance $x$ between the two turbines ($i,j$)  and turbine and terrain specifications that yield the cone opening radius, $r$. The set of upstream turbines, $\mathcal{U}_{id}$ is computed  
by rotating the grid wrt to the current wind regime $d \in \mathcal{D}$ and identifying the turbines, $j$, located upstream of turbine $i$ given $d$. The expression in Eq. (\ref{eq:ss}) is referred to as the \emph{sum-of-squares} (SS) model for 
the total expected power in presence of wakes~\cite{Zhang2014}.  We can now write 
the sum-of-squares (SS) expected power of a wind farm with $m$ turbines:
%~(\cite{Zhang2014}):
\begin{equation}
  E_{SS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \frac{1}{3} u_{id}^3p_d.\label{eq:ess}
\end{equation}

@@

\paragraph{Exact Optimization Methods} 
The WFLO problem can be solved  
to optimality, using
techniques from operations research (e.g., mixed-integer linear programming (MILP), quadratic programming (QP), and constraint programming (CP))~\cite{Zhang2014,turner2014new,donovan2005wind}.
These methods are \emph{exact}, in the sense that they guarantee that the returned solution is indeed
the best solution one can obtain for the given problem. Donovan was the first to solve WFLO using a
MILP approach based on the LS energy expression~\cite{donovan2005wind}. Turner et al.\ \cite{turner2014new} approximated the SS energy expression using a quadratic and a linear approximation. The two resulting methods performed well when
compared to existing approximate and exact techniques.
Another non-linear optimization approach \cite{ulku2019new} relaxed the binary placement variables to continuous values in $[0,1]$
and approximated the constraints such that the resulting variables yield an integer placement.
Their approach provides another approximation to the WFLO problem considered by Turner et al.\ \cite{turner2014new}.

@@

A benchmark study comparing state-of-the-art exact approaches \cite{Zhang2014} to 
approximate methods found that the two perform in a comparable fashion, without large differences between
the attained energy~\cite{yang2019simulated}. Exact methods tend to perform slightly better
in smaller WFLO benchmarks, while simulated annealing methods dominated for larger standard WFLO benchmarks. 
These results lead to the conclusion that both types of methods (exact and approximate) should be considered. Therefore, we introduce a unified paradigm, namely quadratic programming, that enables
both exact and approximate solutions to WFLO. Further, formulating WFLO using quadratic programming enables us to employ 
recent advanced exact software solutions (e.g., Gurobi optimizer) and approximate hardware developments 
(e.g., Fujitsu's Digital Annealer).
%Thus, 
%our approach `enjoys' the best of both worlds (the exact and the approximate).

@@

%% In this section,
%% we formulate the WFLO problem
%% using quadratic programming. Specifically,
We consider two
quadratic optimization problems that represent WFLO: a
quadratic constrained optimization problem (QCOP) and 
a quadratic unconstrained binary optimization (QUBO) problem. 
The former represents WFLO characteristics (proximity and number of turbines)
as hard constraints, while the latter considers these characteristics as soft constraints such that the objective function is penalized in case these
constraints are violated.

@@

To 
maximize $f(x)$ and satisfy the total number of turbines and proximity 
constraints we write
the following quadratic constrained optimization problem (QCOP):
\begin{eqnarray} \label{QCOP}
&\max_{x_i}^{}& \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)x_i x_j)   \\
&s.t.:& \nonumber\\
&\mbox{       }& \sum_{i \in \mathcal{N}}^{} x_i = m,\label{Cardinality}\\
&\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid}\\
&\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}. \label{Vars}
\end{eqnarray} Constraints~\ref{Cardinality}~and~\ref{Grid} ensure that exactly $m$ turbines are placed on the grid
and enforce proximity constraints between the turbines, respectively. We shall refer to the QCOP
formulation of WFLO as \qcls{}.

\qcls{} can be solved using an exact optimization solver, e.g., Gurobi~\cite{gurobi}. 
Linear versions of the QCOP (\qcls) have been previously proposed and solved using the corresponding exact solvers~\cite{Zhang2014,donovan2005wind}.
%\todo{JCB: We need to formally present the ILP model in the experiment section. We can then discuss the fact that the model is identical - because the linearization gives and exact representation. Should put a forward reference to that model here.}
%\todo{Are these linear formulations ``equivalent'' as you claimed or approximations? If they are mathematically equivalent, why do we think a quadratic model will be any better computationally? Arik: The two formulations are equivalent to WFL-QCOP. However, we compare against the ILP model experimentally and it does worse than WFL-QCOP. Hopefully, Jason's part in Section 3.5.1. based on Wen-Yang's thesis will explain why it does better. Originally, I used the quadratic formulation because it seemed more natural. Semantically, the linear model and the quadratic model are the equivalent imho.}

@@

Let $\lambda = (\lambda_1,\lambda_2)$ 
be a vector of constraint weights with $\lambda_i >0$. The equivalent QUBO formulation of the QCOP in Eq. (\ref{QCOP}) is given by,
\begin{equation}\max_{x}^{} f(x) - \lambda_1 (\sum_{i \in \mathcal{N}}^{} x_i -m) ^2 - \lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j , \label{QUBO}\end{equation}
with the term $ \lambda_1 (\sum_{i \in \mathcal{N}}^{} x_i -m) ^2$ 
penalizing solutions that violate the exactly $m$ turbines constraint and the term $\lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j$ penalizing
solutions that violate the proximity constraints. The two penalty terms $\lambda_1$ and $\lambda_2$ 
must be large enough to ensure that the solutions are indeed feasible wrt the problem constraints. We shall refer to the QUBO representation of WFLO as \quls.
%These types of constraints
%are referred to as \emph{soft constraints} as they do not guarantee a feasible solution 
%for an arbitrary selection of $\lambda$.

@@

\begin{mythm}
	The computational complexity of \qcls{} defined in \eqref{QCOP}-\eqref{Vars} is $\mathcal{NP}$-hard. 
\end{mythm}
\begin{myproof}
We prove by showing that a special case of the \qcls{} \eqref{QCOP}-\eqref{Vars}  is
$\mathcal{NP}$-hard. Suppose that $\mathcal{E} = \emptyset$, i.e., the grid is coarse-grained enough to ignore proximity constraints. Then, the constraints in Eq.~(\ref{Grid})
are dropped and we get the heaviest k-subgraph problem, which was proven to be $\mathcal{NP}$-hard~\cite{billionnet2005different}.
Since a special case of \qcls{} is  $\mathcal{NP}$-hard it implies that \qcls{} is at least $\mathcal{NP}$-hard. On the other hand,
\qcls{} is formulated using integer programming (Problem \ref{QCOP}), which means that it is at most $\mathcal{NP}$-hard. Therefore, we get that \qcls{} is $\mathcal{NP}$-hard.
\end{myproof}

@@

In this section, we 
present an empirical evaluation of the two quadratic models based on 
12 standard WFLO instances.
Our main results are: \begin{itemize}
	\item \quls{} when solved using the Digital Annealer (DA) achieves a solution quality as good or better (within time limit) than existing state-of-the-art exact methods on 10 of 12 instances.
		\item To achieve such a solution quality, the runtime for solving \quls{} on the DA is at least two orders of magnitude lower than \qcls{} for hard WFLO instances.  
	\item Given sufficient runtime, solving the \qcls{} model yields strong results on most of the instances. Compared to previous exact models (such as \qcss), \qcls{} is able to find equal or better solutions on 8 of 12 benchmark problems and for 5 of the 6 hard scenarios. %  \todo{Arik: Is this correct? Table 2 shows DA@10seconds with the best solution found for 7/12 instances and \qcls{}grb@3600 for 6/12. And 4/6 vs. 3/6 for the WR36). Are you comparing \qcls{}grb with \qcssgrb and ILP-LS? We seem to be underplaying DA performance in this comment. Chris: I changed to competitive results. You're right, I'd say that the DA is not `losing' to \qcls{}grb.}  
\end{itemize}

@@

Recall that $\mathcal{W}$ is the set of our 12 standard WFLO instances. 
Let $b_w$ be the best energy solution (in terms of SS) obtained for an instance $w \in \mathcal{W}$ 
across the different methods.
%For example, the best solution (within 3600 seconds) for the second instance $w_2$ that aims at placing $30$ turbines on a $10\times10$ grid with a single wind regime is attained by \qulsgrb and the SS energy expression is $15742$ (kW); thus, we set $b_w = 15742$.
Then, for every point in time $t \in \{0,\ldots, 3600\}$ (measured in seconds), we can compute the relative error of a given approach per instance $w$. Let $\mathcal{A}$ be the set of solution approaches, $$\mathcal{A}= \mbox{
	\{\emph{\qulsgrb}}, \mbox{\emph{\qclsgrb}}, \mbox{\emph{\qulsda}}, \mbox{\emph{\ilplsgrb}}, \mbox{\emph{\qcssgrb}}\}.$$ Further, denote by $b_{w,t,a}$ the best solution attained before time $t$ of approach $a$ for instance $w$. Then, the relative error at time $t$ for approach $a$ on instance $w$ is given by
\begin{equation}
  RE(w,t,a) = \frac{b_w - b_{w,t,a}}{b_w}.
\end{equation} Note that the expression is always non-negative, since we are solving a maximization problem. We can now average over the different instances at time $t$. The average performance of approach $a$ at time $t$, $MRE(t,a)$ can be computed as, \begin{equation} MRE(t,a) = \frac{1}{|\mathcal{W}|}\sum_{ w\in\mathcal{W}} RE(w,t,a).\end{equation}
We present $MRE(t,a)$ for each of the approaches to demonstrate the speed of convergence of the proposed LS-based methods.

@@

We observe that the DA-based solution provides
comparable results to state-of-the-art exact methods. For the harder instances (of $36$ wind directions),
the three quadratic LS-based approaches (\qulsda{}, \qulsgrb{}, and \qclsgrb{}) dominate the others by finding the best (or equal) solutions in 5 out of 6 instances. For easier instances, 
the linear approach (\ilplsgrb) finds better solutions.
%\todo{@Chris, Jason: do we have a good explanation for this?}. NO
Table~\ref{tab:results2} shows the same type of results as in Table~\ref{tab:results1}, but now after $3600$ seconds for exact methods. For the exact technique, we can see that the improvement between 100 and 3600 seconds of run-time is marginal. We also observe that the DA at 10 seconds is, therefore, still competitive with the other approaches,
while \qclsgrb{} finds the best solutions in 10 out of the 12 scenarios. 
Our quadratic approaches dominate all six hard scenarios (36 wind directions). %\todo{@Chris: I am not sure that the first table contributes anything beyond the second table. JCB: ``We include the second table to demonstrate that only minor improvements are achieved by the exact techniques between 100 seconds of run-time.''}

We now turn to present the mean relative error over time per approach. Note that lower values indicate better performance.  Figure~\ref{fig:results1} 
presents the MRE over the first $50$ seconds, while Figure~\ref{fig:results2} corresponds
to the full run (after $3600$ seconds).
%The latter provides insights into best performance in terms of energy. The 5 different lines (and their colors) 
%correspond to the 5 WFLO solutions. The horizontal axis represents time in seconds.

\todo{Eldan regarding the first sentence in what follows: I feel like this is not well defined? MRE is a single metric. Do you mean that our approahces have the best MRE both after a short time period (of say 10 seconds) but also at longer time limits (including 3600)?} Our proposed quadratic models, \qcls{} and \quls{},
outperform the rest of the techniques both in terms of runtime (best results in short time), and in terms of overall performance.
Specifically, for most instances, \qclsgrb{} gives the best solution at 3600 seconds, while the DA-based approach
yields competitive solutions after less than 10 seconds. 
Furthermore, 
\qclsgrb{} converges to the best solution faster than all other exact methods. In fact, it is inferior only to the approximate \qulsda{} method.

@@

\begin{sloppypar}To demonstrate the gain in runtime from using the DA, we compare the average speed-up factor between
 \qclsgrb{} and a \qulsda{} solutions after $1$ and $10$ seconds.
For each $w \in \mathcal{W}$ we computed the mean $S(\mbox{\emph{\qclsgrb}}, \mbox{\emph{\qulsda}}, t, w)$ for $t \in \{1,10\}$. 
%and $S(\qclsgrb, \qulsda, 10sec, w)$
%for every $w$ 
%and averaged the speed-up factors across the instances. 
The average speed-up factors were found to be $666$ and $100$ for $1$ and $10$ seconds, respectively. 
Per instance, the \emph{minimum} speedup was $1$ for $1$ second and $0.1$ for $10$ seconds: for some instances \qclsgrb{} achieved the same quality as \qulsda{} in less time. The \emph{maximum} speedup was (at least) $3600$ for $1$ second and (at least) $360$ for $10$ seconds: for those instances
\qclsgrb{} was not able to find a comparable solution to \qulsda{} within the $3600$ time limit. As a consequence, 
the \qulsda{} speed-up is actually even higher than computed above. \end{sloppypar}
%\footnote{The speed-up factors  were rounded to have 0 values after the floating point.}

%These results show that the DA achieves competitive solutions after running for $1$ second.
%Many of the instances running \qclsgrb for 3600 seconds did not result in as good a solution as \qulsda at time $t$ (1 or 10 seconds).
%we consider the speed-up factor to be a lower bound, meaning
When compared
to the other LS-based methods, the DA speed-up is even larger than when compared to \qclsgrb{}.  At $1$ ($10$) seconds, the mean speedup factor compared to \ilplsgrb{} 
is $1385$ ($180$) and for the \qulsgrb{} it is $900$ ($152$). We compare against the LS based methods as they directly optimize on the same objective as the DA.  
For full speedup results see Tables~\ref{tab:results3} and~\ref{tab:results4} in~\ref{app1}.
%speed-up factor, which we did not compute due to the time limit of $3600$. 
%This is an encouraging result, especially due to the pessimistic computational complexity of WFLO.
